
% File: executive_summary.tex

\newpage
\thispagestyle{plain}

\addcontentsline{toc}{section}{Executive Summary}

\begin{center}
\textbf{\LARGE Executive Summary}
\end{center}

\vspace{1cm}

\begin{spacing}{1.5}

Affordable housing remains one of the most pressing challenges faced by students and young professionals across Pakistan. The traditional rental process is dominated by informal brokers, unverified listings, and a complete absence of any structured method for finding a compatible roommate. As a result, tenants frequently encounter inflated brokerage commissions, fraudulent advertisements, security risks, and stressful living arrangements caused by mismatched roommates. Existing online platforms are oriented toward full-house rentals and provide no support for single rooms, bed spaces, or shared accommodation located near universities and workplaces.

This project, titled \textbf{Property Rental System}, proposes a unified, web-based ecosystem that directly addresses these gaps. Property Rental System combines a dedicated micro-rental marketplace for single rooms, bed spaces, and hostels with an intelligent, algorithm-driven roommate matching engine that scores potential roommates on the basis of lifestyle, preferences, budget, location, and daily schedule compatibility. To establish trust, the platform enforces a multi-layer verification system covering email and phone confirmation, administrator-reviewed CNIC verification, and property ownership validation. A commute-based search powered by the Google Maps API enables users to discover accommodation within walking distance of their university or office, while secure in-app messaging, structured booking requests, online payment recording, and auto-generated digital rental agreements automate the entire rental lifecycle.

Property Rental System is implemented using the MERN technology stack, comprising MongoDB, Express.js, React, and Node.js, with TypeScript applied across both the client and server for type safety. The system follows a layered, role-based architecture supporting three principal user classes: Tenant, Landlord, and Administrator. Geospatial querying, JSON Web Token authentication, role-based access control, cloud-based media storage, and automated PDF generation form the technical backbone of the platform.

The expected contribution of Property Rental System is a safer, more affordable, and more transparent rental experience for an underserved segment of the market. By reducing dependence on informal brokers, mitigating the risk of fraud, and replacing the roommate ``gamble'' with data-driven compatibility matching, the platform delivers measurable social value while remaining commercially viable. This document presents the complete software requirements specification and software design for the system, providing the formal foundation for its implementation, testing, and evaluation.

\end{spacing}
\newpage
